### 2017-04-10 paper/oopsla17-submission/gallery.tex
use colors instaed of bold/underline

Next, we present a \emph{qualitative} evaluation that compares the
predictions made by our classifiers with those of \sherrloc.
%
In particular, we demonstrate, with a series of example programs from
our student dataset, how our classifiers are able to use past student
mistakes to make more accurate predictions of future fixes.
%
For each example, we provide
%
(1) the code,
%
(2) \sherrloc's prediction (in \textbf{bold}),
%
(3) the prediction of our \dectree classifier (\underline{underlined}), and
%
(4) an \emph{explanation} of why our classifier made its prediction.
%
We choose the \dectree classifier for this section as its model is more
easily interpreted than the MLP.\@
%
We also exclude the \ExprSize feature from the model used in this
section, as it makes the predictions harder to motivate, and as we saw
in \autoref{sec:feature-utility} it does not appear to contribute
significantly to the model's accuracy.
%
We explain the predictions by analyzing the paths induced in
the decision tree by the features of the input programs.
%
% We will also attempt, for each program, to explain \emph{why} the
% Decision Tree made its prediction, by analyzing the paths induced
% by the programs.

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
We begin with a few programs where our classifier fails to
make the correct prediction.
%
For these programs we will additionally \colorbox{red!50}{highlight} the
correct blame location.

\paragraph{Constructing a List of Duplicates}
% data/sp14/0219.ml
Our first program is a simple recursive function |clone| that takes an
item |x| and a count |n|, and produces a list containing |n| copies of
|x|.
%
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
    loop [] n
\end{ecode}
% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
%
The student has defined a helper function |loop| with an accumulator
|acc|, likely meant to call itself tail-recursively.
%
Unfortunately, she has called the top-level function |clone| rather than
|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
for the |x| argument to |clone|.

@@

\paragraph{Currying Considered Harmful?}
% data/sp14/1204.ml
Our next example is another ill-fated attempt at |clone|.
%
\begin{ecode}
  let rec clone x n =
    let rec loop (*@\colorbox{red!50}{x n acc}@*) =
      if n < 0 then
        acc
      else
        ==loop== __(x, (n - 1), (x :: acc))__ in
    loop (x, n, [])
\end{ecode}
The issue here is that \ocaml functions are \emph{curried} by default
--- \ie they take their arguments one at a time --- but our student has
called the inner |loop| with all three arguments in a tuple.
%
Many experienced functional programmers would choose to keep |loop|
curried and rewrite the calls, however our student decides instead to
\emph{uncurry} |loop|, making it take a tuple of arguments.
%
\sherrloc blames the recursive call to |loop| while our classifier
blames the tuple of arguments --- a reasonable suggestion, but alas not
the answer the student was looking for.

@@

\paragraph{Extracting the Digits of an Integer}
% data/sp14/2176.ml
Consider first a simple recursive function |digitsOfInt| that extracts
the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ __[ ___=digitsOfInt (n / 10)=___ ]__
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
actually correct!

@@

The student first defines a helper function |clone|, as seen above.
%
\begin{ecode}
  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)
\end{ecode}
%
Then she defines |padZero| with a simple branch that determines which
list is shorter, followed by a |clone| for the appropriate number of
zeros.
%
\lstset{firstnumber=last}
\begin{ecode}
  let padZero l1 l2 =
    let n = List.length l1 - List.length l2 in
    if n < 0 then
      (clone 0 ((-1) * n) @ l2, l2)
    else
      (l1, _=clone 0 n=___ :: l2__)
\end{ecode}
\lstset{firstnumber=1}
%
Alas, our student has accidentally used the |::| operator rather than
the |@| operator in the |else| branch.
%
\sherrloc and \ocaml correctly determine that she cannot cons the
|int list| returned by |clone| onto |l2|, which is another |int list|,
but they decide to blame the call to |clone|, while our
\dectree correctly blames the |::| constructor.

>>>>

\definecolor{tree}{HTML}{66C2A5}
\definecolor{sherrloc}{HTML}{FC8D62}
\definecolor{fix}{HTML}{8DA0CB}

Next, we present a \emph{qualitative} evaluation that compares the
predictions made by our classifiers with those of \sherrloc.
%
In particular, we demonstrate, with a series of example programs from
our student dataset, how our classifiers are able to use past student
mistakes to make more accurate predictions of future fixes.
%
For each example, we provide
%
(1) the code,
%
(2) \colorbox{sherrloc!75}{\sherrloc's} prediction,
%
(3) our \colorbox{tree!75}{\dectree's} prediction, and
%
(4) an \emph{explanation} of why our classifier made its prediction.
%
We choose the \dectree classifier for this section as its model is more
easily interpreted than the MLP.\@
%
We also exclude the \ExprSize feature from the model used in this
section, as it makes the predictions harder to motivate, and as we saw
in \autoref{sec:feature-utility} it does not appear to contribute
significantly to the model's accuracy.
%
We explain the predictions by analyzing the paths induced in
the decision tree by the features of the input programs.
%
% We will also attempt, for each program, to explain \emph{why} the
% Decision Tree made its prediction, by analyzing the paths induced
% by the programs.

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
We begin with a few programs where our classifier fails to
make the correct prediction.
%
For these programs we will additionally \colorbox{fix!75}{highlight} the
correct blame location.

\paragraph{Constructing a List of Duplicates}
% data/sp14/0219.ml
Our first program is a simple recursive function |clone| that takes an
item |x| and a count |n|, and produces a list containing |n| copies of
|x|.
%
% \begin{ecode}
%   let rec clone x n =
%     let loop acc n =
%       if n <= 0 then
%         acc
%       else
%         (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
%     loop [] n
% \end{ecode}
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        (*@\colorbox{fix!75}{clone} \colorbox{sherrloc!75}{([\colorbox{tree!75}{x}] @\ acc)}@*) (n - 1) in
    loop [] n
\end{ecode}
% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
%
The student has defined a helper function |loop| with an accumulator
|acc|, likely meant to call itself tail-recursively.
%
Unfortunately, she has called the top-level function |clone| rather than
|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
for the |x| argument to |clone|.

@@

\paragraph{Currying Considered Harmful?}
% data/sp14/1204.ml
Our next example is another ill-fated attempt at |clone|.
%
\begin{ecode}
  let rec clone x n =
    let rec loop (*@\colorbox{fix!75}{x n acc}@*) =
      if n < 0 then
        acc
      else
        (*@\colorbox{sherrloc!75}{loop} \colorbox{tree!75}{(x, (n - 1), (x :: acc))}@*) in
    loop (x, n, [])
\end{ecode}
%
The issue here is that \ocaml functions are \emph{curried} by default
--- \ie they take their arguments one at a time --- but our student has
called the inner |loop| with all three arguments in a tuple.
%
Many experienced functional programmers would choose to keep |loop|
curried and rewrite the calls, however our student decides instead to
\emph{uncurry} |loop|, making it take a tuple of arguments.
%
\sherrloc blames the recursive call to |loop| while our classifier
blames the tuple of arguments --- a reasonable suggestion, but alas not
the answer the student was looking for.

@@

\paragraph{Extracting the Digits of an Integer}
% data/sp14/2176.ml
Consider first a simple recursive function |digitsOfInt| that extracts
the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ (*@\colorbox{tree!75}{[ \colorbox{sherrloc!75}{digitsOfInt (n / 10)} ]}@*)
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
actually correct!

@@

The student first defines a helper function |clone|, as seen above.
%
\begin{ecode}
  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)
\end{ecode}
%
Then she defines |padZero| with a simple branch that determines which
list is shorter, followed by a |clone| for the appropriate number of
zeros.
%
\lstset{firstnumber=last}
\begin{ecode}
  let padZero l1 l2 =
    let n = List.length l1 - List.length l2 in
    if n < 0 then
      (clone 0 ((-1) * n) @ l2, l2)
    else
      (l1, (*@\colorbox{tree!75}{\colorbox{sherrloc!75}{clone 0 n} :: l2}@*))
\end{ecode}
\lstset{firstnumber=1}
%
Alas, our student has accidentally used the |::| operator rather than
the |@| operator in the |else| branch.
%
\sherrloc and \ocaml correctly determine that she cannot cons the
|int list| returned by |clone| onto |l2|, which is another |int list|,
but they decide to blame the call to |clone|, while our
\dectree correctly blames the |::| constructor.
